---
title: "A potato Analysis"
output: html_document
date: "2024-03-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### *Introduction*


**Goal**: You should model the **Flavor score** as a function of the explanatory factors
(i.e. leave out Texture score and Moistness score)

**Dataset**:  potato.dat

**Source**: A. Mackey and J. Stockman (1958). "Cooking Quality of Oregon-Grown
Russet Potatoes", American Potato Journal, Vol.35, pp395-407

**Description**: Texture, Flavor, and Moistness Scores for potatoes
of 2 sizes, from two areas, two holding temps, 4 holding periods,
and 5 cooking methods. Each score is an average of 20 scores 
(5 judges x 4 replications).

Variables/Columns

Growing Area   8  /*  1=Southern OR,  2=Central OR  */
Two week holding temp   16   /* 1=75F   2=40F   */
Size of Potato    24   /* 1=Large, 2=Medium  */
Storage Period  32  /* 1=0 months, 2=2 mths 3=4mths,  4=6mths */
Cooking method  40 /* 1=Boil, 2=Steam, 3=Mash, 4=Bake@350, 5=Bake@450  */
Texture score  46-48
Flavor score   54-56
Moistness score  62-64



#### *Importing the dataset from URL*

```{r} 

url <- "http://users.stat.ufl.edu/~winner/data/potato.dat"
potato_data <- read.table(url, header = FALSE)

colnames(potato_data) <- c("Growing_Area", "Temperature", "Size", "Storage_period", "Cooking_Method", "Texture", "Flavor", "Moistness")

potato_data <- potato_data[, !colnames(potato_data) %in% c("Texture", "Moistness")]

#Transforming into factors 

potato_data$Growing_Area <- factor(potato_data$Growing_Area)
potato_data$Temperature <- factor(potato_data$Temperature)
potato_data$Size <- factor(potato_data$Size)

potato_data$Storage_period <- factor(potato_data$Storage_period, levels = c(1, 2, 3, 4),  ordered = TRUE) #set storage period to different levels, ordered because time period

levels(potato_data$Storage_period) <- c("0", "2", "4", "6")
levels(potato_data$Cooking_Method) <- c("Boil", "Steam", "Mash", "Bake_350", "Bake_450")

potato_data$Cooking_Method <- factor(potato_data$Cooking_Method, levels = c(1, 2, 3, 4,5))
#set cooking method to different levels, not ordered


potato_data



```
#### *Pre analysis to choose the regression model*


```{r}


#overall mean
overall_mean_flavor <- mean(potato_data$Flavor)
cat("Mean flavor score overall:", overall_mean_flavor, "\n")

#Compute the mean flavor score for each different area
mean_flavor_area <- tapply(potato_data$Flavor, potato_data$Growing_Area, mean)
cat("Mean flavor score for each area:\n")
print(mean_flavor_area)

# 3. Compute the mean flavor score for each temperature
mean_flavor_temperature <- tapply(potato_data$Flavor, potato_data$Temperature, mean)
cat("Mean flavor score for each temperature:\n")
print(mean_flavor_temperature)


#overall standard deviation:

sd_flavor_overall <- sd(potato_data$Flavor)
cat("Standard deviation flavor score overall:", sd_flavor_overall, "\n")

#Compute the sd flavor score for each different area
sd_flavor_area <- tapply(potato_data$Flavor, potato_data$Growing_Area, sd)
cat("Standard deviation flavor score for each area:\n")
print(sd_flavor_area)

# 3. Compute the mean flavor score for each temperature
sd_flavor_temperature <- tapply(potato_data$Flavor, potato_data$Temperature, sd)
cat("Standard deviation flavor score for each temperature:\n")
print(sd_flavor_temperature)


#Some boxplots (we only have discrete variables except the Flavor Score)

layout(matrix(1:6, nrow = 2, byrow = TRUE)) #permet d'afficher les prochains plots dans un "tableau" 1 row, 2 columns)

boxplot(Flavor ~ Growing_Area, data = potato_data, names = c("Southern", "Central"), ylab = "Flavor Score", xlab = "Growing Area")
abline(h = overall_mean_flavor, col = "red", lty = 2)

boxplot(Flavor ~ Temperature, data = potato_data, names = c("75F", "40F"), ylab = "Flavor Score", xlab = "Holding Temperature")
abline(h = overall_mean_flavor, col = "red", lty = 2)

boxplot(Flavor ~ Size, data = potato_data, names = c("medium", "big"), ylab = "Flavor Score", xlab = "Size")
abline(h = overall_mean_flavor, col = "red", lty = 2)

boxplot(Flavor ~ Storage_period, data = potato_data, names = c("0", "2", "4", "6"), ylab = "Flavor Score", xlab = "Storage Period (in months)")
abline(h = overall_mean_flavor, col = "red", lty = 2)

overall_median_flavor <- median(potato_data$Flavor)

cooking_methods_names = c("Boil", "Steam", "Mash", "Bake 350F", "Bake 450F")

boxplot(Flavor ~ Cooking_Method, data = potato_data, names = cooking_methods_names, ylab = "Flavor Score", xlab = "Cooking Method")
abline(h = overall_median_flavor, col = "red", lty = 2)

layout(matrix(1:1, nrow = 1, byrow = TRUE)) #Retour à un seul plot



interaction.plot(potato_data$Storage_period, potato_data$Cooking_Method, potato_data$Flavor,trace.label = "Cooking Method", xlab = "Storage Period (Month)", ylab = "Flavor Score")
```
Seems that the size of the potato does not affect the Flavor score (mean doesn't change much). We suspect that there is interaction between the Storage Period and the cooking method, so we might want to include it into our regression Model.

#### *Fitting an ANOVA model*

In this case, the regression model has the formula:

<p style="text-align: center;"> Weightgain = $\beta_0$ + $\beta_1 \cdot$ sourceCereal + $\beta_2 \cdot$ typeLow +$\beta_3 \cdot$ sourceCereal $\cdot$ typeLow</p>

where the covariate are indicator variables.

```{r}


flavor.aov <- aov(Flavor ~ Storage_period*Cooking_Method, data=potato_data) 

flavor_1.aov <- aov(Flavor ~ Growing_Area + Cooking_Method*Storage_period + Temperature*Storage_period + Cooking_Method*Temperature, data=potato_data) 

summary(flavor_1.aov)

coef(flavor_1.aov)

layout(matrix(1:6,ncol=3)) 
plot(flavor_1.aov, which=c(1,2,3,4,5,6))



```

Checker lab 2 pour interprétation des plots de fin

