---
title: "Potato_analysis_raph"
output: html_document
date: "2024-03-20"
---

# Load potato data
```{r}
# Read text file
data <- read.table("http://users.stat.ufl.edu/~winner/data/potato.dat", header = FALSE)

# Assign column names
colnames(data) <- c("Growing_Area", "Temperature", "Size", "Storage_period", "Cooking_Method", "Texture", "Flavor", "Moistness")

# Visualize data
data
```
# Boxplots
```{r}
# Create boxplots for each variable based on flavor
layout(matrix(1:6, nrow = 2, byrow = TRUE))

boxplot(Flavor ~ Growing_Area, data = data, names = c("Southern", "Central"), ylab = "Flavor Score", xlab = "Growing Area")
abline(h = mean(data$Flavor), col = "red", lty = 2)

boxplot(Flavor ~ Temperature, data = data, names = c("75F", "40F"), ylab = "Flavor Score", xlab = "Holding Temperature")
abline(h = mean(data$Flavor), col = "red", lty = 2)

boxplot(Flavor ~ Size, data = data, names = c("medium", "big"), ylab = "Flavor Score", xlab = "Size")
abline(h = mean(data$Flavor), col = "red", lty = 2)

boxplot(Flavor ~ Storage_period, data = data, names = c("0", "2", "4", "6"), ylab = "Flavor Score", xlab = "Storage Period (in months)")
abline(h = mean(data$Flavor), col = "red", lty = 2)

boxplot(Flavor ~ Cooking_Method, data = data, names = c("Boil", "Steam", "Mash", "Bake 350F", "Bake 450F"), ylab = "Flavor Score", xlab = "Cooking Method")
abline(h = mean(data$Flavor), col = "red", lty = 2)

```
Based on the boxplots, we observe the most difference in means for the Storage period and the Cooking method.

# Fit ANOVA model
```{r} 
# ANOVA model with 2 predictors
storage.aov <- aov( Flavor ~ Storage_period, data=data)
sum.aov <- aov( Flavor ~ Storage_period+Cooking_Method, data=data)
product.aov <- aov( Flavor ~ Storage_period*Cooking_Method, data=data)

# Summary and model coefficients
summary(storage.aov)
coef(storage.aov)

summary(sum.aov)
coef(sum.aov)

summary(product.aov)
coef(product.aov)

```
# Interpretation of model coefficients
For these models, the Cooking method doesn't have a significant p-value. We can't reject the null hypothesis, which means the difference in means observed can't be attributed to a significant relationship between the cooking method and the flavor.

However, the Storage period is significant. To have a better idea whether a simple model using this variable would be sufficient or if other predictors would make a better model, we can try using stepwise selection.

```{r}
step <- step(aov( Flavor ~ Storage_period + Cooking_Method + Growing_Area + Temperature + Size, data=data))
```
From this selection the best model would be modelling the Flavor as a sum if the Storage_period, Growing_Area and Temperature
```{r}
best.aov <- aov( Flavor ~ Storage_period + Growing_Area + Temperature, data=data)

summary(best.aov)
coef(best.aov)
```
Indeed, all predictors used show significant relationship to the Flavor.

# Model assessment
To assess the validity of the model, we observe the residuals and QQ plots
```{r}
# Residuals and QQ plots
layout(matrix(1:4,ncol=2))
plot(best.aov)
```
QQ plot shows mostly a linear relationship between standardized and QQ residuals, with some differences at the extremities but reasonably enough to assume gaussianity and homoscedasticity in this model.